Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-14 paper.tex
consolidate analysis into single script. Add no fuel shortage build schedule and a bit of analysis
added direct separated pu inv comparison
paragraph about pu inventory comparison fig

A few other small changes to the scenario are also necessary.  For the cases
using quarterly time steps with individual reactors, cycle and outage times
must be adjusted to be a multiple of 3 months.  In order to make the
comparison fair, the cases with monthly time steps are also adjusted to have
identical cycle and outage times.  The invariants preserved with respect to
reactor behavior are shown in Table \ref{tab:exp2-invar}.  Table
\ref{tab:exp2-reactors} shows the computed/selected configuration for both
reactor types.  In each of the four cases, the initial 100 LWRs are modified
to have a zero length refueling outage with an explicit 900 MWe power
capacity.  This was done to avoid having all the initial LWRs in the
individual reactor cases refuel at the same time.

@@

    \label{tab:exp2-invar}
\end{table}

@@

    \label{tab:exp2-reactors}
\end{table*}

@@

Figure \ref{fig:exp2-power} shows an overall view of the generated
power over time for each of the four cases.  Figure
\ref{fig:exp2-power-rel} normalizes the Figure
\ref{fig:exp2-power} curves to the expected exponential power curve
magnifying some interesting differences between the four cases.  For the first
100 years, the four cases behave somewhat similarly, although the individually
modeled reactors in cases 1 and 3 actually go online and offline for refueling
causing more variance.  Around year 100, a fuel shortage begins and more
significant differences between the four cases become apparent. This fuel
shortage lasts until about year 140. The power generated in Case 4 is lower
than in case 2 during the fuel shortage years. This is primarily due to the
\emph{quantized shutdown effect}.  The longer 3-month time step in case 4
results in reactor capacity going offline longer than necessary.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/power.eps}
    \caption[Generated power]{
        Generated power for all four cases.  A fuel shortage occurs from about
        year 100 to about year 140.  In cases 1 and 3 with individual reactor
        modeling, refueling cycles become much more synchronized during and
        after the fuel shortage resuting in large power fluctuations (the
        \emph{cycle staggering effect}).
    }
    \label{fig:exp2-power}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/power-rel.eps}
    \caption[Normalized power]{
        Generated power is normalized to the expected 1\% exponential growth
        curve for all four cases.  A fuel shortage occurs from about year 100
        to about year 140.  The \emph{cycle staggering effect} can be seen as
        divergence from 1.0 and is especially visible in cases 1 and 3 where
        the shortage increases cycle synchronization significantly. The
        \emph{quantized shutdown effect} can also be seen in the discrepency
        between quarterly and monthly time steps for fleet reactor modeling
        (cases 2 and 4) during the shortage; a longer time step causes more
        reactor outage than necessary. The consistent jumping up and down for
        the fleet reactor cases is due to the mis-alignment of reactors going
        offline (multiples of 12 months) and the build period (21 months).
    }
    \label{fig:exp2-power-rel}
\end{figure*}

@@

Figure \ref{fig:exp2-unfueled} shows the fuel shortage more explicitly
for each of the four cases. The individual reactor modeling cases result in
more cumulative outage than corresponding fleet-based cases. The quarterly
time step of cases 3 and 4 make the outage significantly more severe. Cases 1
and 3 exhibit approximately twice as much offline power as their corresponding
fleet based simulations in cases 2 and 4.  The monthly time step simulations
also have roughly twice as much offline power as their corresponding quarterly
cases.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/unfueled.eps}
    \caption[Cumulative offline power due to fuel shortage]{
        Cumulative offline power for reactors that had delayed cycle start due
        to fuel shortage. The difference between individual and fleet based
        modeling is primarily a result of the \emph{fuel sharing effect}. The
        difference between monthly and quarterly time steps is primarily
        caused by the \emph{inventory drawdown effect}.
    }
    \label{fig:exp2-unfueled}
\end{figure}

Quantifying imperfect fuel sharing is a bit difficult, but Figure
\ref{fig:exp2-badshare} gives one way to approximate the \emph{fuel
sharing effect}.  Fleet reactor modeling by design exhibits perfect fuel
sharing by design (zero inefficiency) and so is not included in this figure.
Every batch that is given to a reactor that ends up not being able to start it
cycle at the scheduled time (due to fuel shortage) adds one to this cumulative
for each time step the cycle start is delayed.

One useful over-estimate of the fuel sharing inefficiency is to assume every
one of these batches could have been given to a reactor enabling it to
operate.  Multiplying each of these cumulative batches by reactor capacity
(i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
overall fuel outage.  Correcting for capacity factor in is appropriate these
cases because the outage is many reactor cycle lengths in duration. This
results in a cumulative offline power caused by poor fuel sharing of roughly
1,600 $\text{GWe} \cdot \text{months}$ and 5,400 $\text{GWe} \cdot
\text{months}$ for cases 1 and 3 respectively.  These approximations are a
significant portion of the difference between individual and fleet cases in
Figure \ref{fig:exp2-unfueled} suggesting that most of the discrepency
in cumulative outage between fleet and individual modeling is caused by
inefficient fuel sharing in the individual reactor cases.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/badshare.eps}
    \caption[Cumulative unnecessary idling fuel]{
        Fuel sharing inefficiency approximated by the cumulative number of
        fuel batches given to each reactor on each refueling period multiplied
        by the number of time steps that reactor had to delay the start of its
        next cycle (due to fuel shortage).  Fleet reactors have perfect fuel
        sharing by design and so are not included here. The large difference
        between cases 1 and 3 is mostly due to the \emph{inventory drawdown
        effect}. 
    }
    \label{fig:exp2-badshare}
\end{figure}

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:exp2-unfueled} and \ref{fig:exp2-badshare} are
primarily caused by the \emph{inventory drawdown effect}.  Figure
\ref{fig:exp2-puinv} helps to visualize this effect.  The black dots in
the figure represent impulse flows of available Pu out of the fuel fab and
into fast reactors.  Blue lines represent the flows into Pu inventory
available for making fuel. The in flows are not shown for the individual
reactor cases because they are somewhat messy and obscure otherinformation in
the plots. The colored curves show Pu inventory available for making fast
reactor fuel.  When the inventory curves meet and force down the flows, fuel
shortages occur.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv.eps}
    \caption[Pu inventory and flow]{
        Pu inventory and flows for all four cases.  In-flows for cases 1 and 3
        are omitted because they are very noisy and obscure useful other
        information.  Larger time steps results in larger per time step
        material flows. These larger flows mean more material is not available
        for supplying to requesters.  This information lag causes more supply
        constraints than occur with smaller time steps - illustrating the
        \emph{inventory drawdown effect}.
    }
    \label{fig:exp2-puinv}
\end{figure*}

@@

The impact of modeling assumptions depends heavily on the metrics of interest.
For example, someone interested in fuel-shortage driven reactor outages could
observe a factor of four span among the four cases shown in the detail
comparison.  However, someone more interested in the overall transition speed
would not see as significant a discrepancy.

>>>>

A few other small changes to the scenario are also necessary.  For the cases
using quarterly time steps with individual reactors, cycle and outage times
must be adjusted to be a multiple of 3 months.  In order to make the
comparison fair, the cases with monthly time steps are also adjusted to have
identical cycle and outage times.  The invariants preserved with respect to
reactor behavior are shown in Table \ref{tab:invar}.  Table
\ref{tab:reactors} shows the computed/selected configuration for both
reactor types.  In each of the four cases, the initial 100 LWRs are modified
to have a zero length refueling outage with an explicit 900 MWe power
capacity.  This was done to avoid having all the initial LWRs in the
individual reactor cases refuel at the same time.

@@

    \label{tab:invar}
\end{table}

@@

    \label{tab:reactors}
\end{table*}

@@

Figure \ref{fig:power} shows an overall view of the generated
power over time for each of the four cases.  Figure
\ref{fig:power-rel} normalizes the Figure
\ref{fig:power} curves to the expected exponential power curve
magnifying some interesting differences between the four cases.  For the first
100 years, the four cases behave somewhat similarly, although the individually
modeled reactors in cases 1 and 3 actually go online and offline for refueling
causing more variance.  Around year 100, a fuel shortage begins and more
significant differences between the four cases become apparent. This fuel
shortage lasts until about year 140. The power generated in Case 4 is lower
than in case 2 during the fuel shortage years. This is primarily due to the
\emph{quantized shutdown effect}.  The longer 3-month time step in case 4
results in reactor capacity going offline longer than necessary.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/power.eps}
    \caption[Generated power]{
        Generated power for all four cases.  A fuel shortage occurs from about
        year 100 to about year 140.  In cases 1 and 3 with individual reactor
        modeling, refueling cycles become much more synchronized during and
        after the fuel shortage resuting in large power fluctuations (the
        \emph{cycle staggering effect}).
    }
    \label{fig:power}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/power-rel.eps}
    \caption[Normalized power]{
        Generated power is normalized to the expected 1\% exponential growth
        curve for all four cases.  A fuel shortage occurs from about year 100
        to about year 140.  The \emph{cycle staggering effect} can be seen as
        divergence from 1.0 and is especially visible in cases 1 and 3 where
        the shortage increases cycle synchronization significantly. The
        \emph{quantized shutdown effect} can also be seen in the discrepency
        between quarterly and monthly time steps for fleet reactor modeling
        (cases 2 and 4) during the shortage; a longer time step causes more
        reactor outage than necessary. The consistent jumping up and down for
        the fleet reactor cases is due to the mis-alignment of reactors going
        offline (multiples of 12 months) and the build period (21 months).
    }
    \label{fig:power-rel}
\end{figure*}

@@

Figure \ref{fig:unfueled} shows the fuel shortage more explicitly
for each of the four cases. The individual reactor modeling cases result in
more cumulative outage than corresponding fleet-based cases. The quarterly
time step of cases 3 and 4 make the outage significantly more severe. Cases 1
and 3 exhibit approximately twice as much offline power as their corresponding
fleet based simulations in cases 2 and 4.  The monthly time step simulations
also have roughly twice as much offline power as their corresponding quarterly
cases.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/unfueled.eps}
    \caption[Cumulative offline power due to fuel shortage]{
        Cumulative offline power for reactors that had delayed cycle start due
        to fuel shortage. The difference between individual and fleet based
        modeling is primarily a result of the \emph{fuel sharing effect}. The
        difference between monthly and quarterly time steps is primarily
        caused by the \emph{inventory drawdown effect}.
    }
    \label{fig:unfueled}
\end{figure}

Quantifying imperfect fuel sharing is a bit difficult, but Figure
\ref{fig:badshare} gives one way to approximate the \emph{fuel
sharing effect}.  Fleet reactor modeling by design exhibits perfect fuel
sharing by design (zero inefficiency) and so is not included in this figure.
Every batch that is given to a reactor that ends up not being able to start it
cycle at the scheduled time (due to fuel shortage) adds one to this cumulative
for each time step the cycle start is delayed.

One useful over-estimate of the fuel sharing inefficiency is to assume every
one of these batches could have been given to a reactor enabling it to
operate.  Multiplying each of these cumulative batches by reactor capacity
(i.e. 450 MWe) provides a way to compare fuel sharing inefficiency with the
overall fuel outage.  Correcting for capacity factor in is appropriate these
cases because the outage is many reactor cycle lengths in duration. This
results in a cumulative offline power caused by poor fuel sharing of roughly
1,600 $\text{GWe} \cdot \text{months}$ and 5,400 $\text{GWe} \cdot
\text{months}$ for cases 1 and 3 respectively.  These approximations are a
significant portion of the difference between individual and fleet cases in
Figure \ref{fig:unfueled} suggesting that most of the discrepency
in cumulative outage between fleet and individual modeling is caused by
inefficient fuel sharing in the individual reactor cases.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.0\columnwidth]{exp2/badshare.eps}
    \caption[Cumulative unnecessary idling fuel]{
        Fuel sharing inefficiency approximated by the cumulative number of
        fuel batches given to each reactor on each refueling period multiplied
        by the number of time steps that reactor had to delay the start of its
        next cycle (due to fuel shortage).  Fleet reactors have perfect fuel
        sharing by design and so are not included here. The large difference
        between cases 1 and 3 is mostly due to the \emph{inventory drawdown
        effect}. 
    }
    \label{fig:badshare}
\end{figure}

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are
primarily caused by the \emph{inventory drawdown effect}.  Figure
\ref{fig:puinv} helps to visualize this effect.  The black dots in
the figure represent impulse flows of available Pu out of the fuel fab and
into fast reactors.  Blue lines represent the flows into Pu inventory
available for making fuel. The in flows are not shown for the individual
reactor cases because they are somewhat messy and obscure otherinformation in
the plots. The colored curves show Pu inventory available for making fast
reactor fuel.  When the inventory curves meet and force down the flows, fuel
shortages occur.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv.eps}
    \caption[Pu inventory and flow]{
        Pu inventory and flows for all four cases.  In-flows for cases 1 and 3
        are omitted because they are very noisy and obscure useful other
        information.  Larger time steps results in larger per time step
        material flows. These larger flows mean more material is not available
        for supplying to requesters.  This information lag causes more supply
        constraints than occur with smaller time steps - illustrating the
        \emph{inventory drawdown effect}.
    }
    \label{fig:puinv}
\end{figure*}

@@

The impact of modeling assumptions depends heavily on the metrics of interest.
For example, someone interested in fuel-shortage driven reactor outages could
observe a factor of four span among the four cases shown in the detail
comparison.  However, someone more interested in the overall transition speed
would not see as significant a discrepancy.  Also notable is that these
different modeling choices may have varying levels of significance depending
on the scenario/context they are operating in.  Figure \ref{fig:puinv-compare}
shows the same inventory curves from Figure \ref{fig:puinv} superimposed.
Before the fuel shortage begins, differences between each of the four cases
are relatively minor.  However, after the shortage begins around year 100,
separated Pu inventories vary much more significantly between the different
cases.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/puinv-compare.eps}
    \caption[TODO]{
        TODO
    }
    \label{fig:puinv-compare}
\end{figure*}
